Draft final report\\ Measles risk assessment, modelling and benefit--cost analysis\\ \vspace{2 mm} {\large David T S Hayman, Tim Carpenter,\\ Jonathan C Marshall, Mick Roberts, Nigel P French}

### 2014-10-29 draftfinalreport.tex
revisions

The analyses we believe could further inform the understanding of risk from measles infection are:
\begin{itemize}
\item Multivariate modelling to account for confounding within the univariate analyses.
\item Inclusion of multivariate model results with local level immunity data to develop a risk map for New Zealand.
\item Analyses of changing risk factors through time during outbreaks.
\item Modelling transmission chains in the population to understand effective reproduction number.
\item Development of risk maps to understand measles importation.
\end{itemize}

\subsection{Additional risk analyses}
\label{sub:risk_analyses}

In this section we provide a brief description of the work that we believe will help inform the Ministry of Health regarding the understanding of risk from measles. These analyses are intended to build on the analyses already included in the \emph {Progress Towards Measles Elimination in New Zealand - Final} report reviewed above.

@@

In light of the apparent increasing trend in measles incidence in the last few years (Figure~\ref{fig:incidence1997}), we reviewed the information on measles importation and the origins of the introductions of measles into New Zealand. To help understand the risk of measles importation, with a particular goal of enabling the Ministry to better inform travellers and understand high risk periods, we sought to use quantitative methods to evaluate the risk of measles importation.

\subsection{Additional risk analyses methods}

For our initial analyses, we use arrivals data from New Zealand immigration (\href{http://www.immigration.govt.nz/}{www.immigration.govt.nz}) as a proxy for human movement to and from New Zealand. We collated country population size, measles incidence and measles vaccination cover from the WHO (\href{http://www.who.int/research/en/}{www.who.int/research/en/}). Note this uses all immigration of foreign nationals, coming for whatever purpose. We used the WHO data to determine per capita measles cases for each year and used these data and the number of immigrants to New Zealand to begin to understand where measles was likely to be imported from.

\begin{table}
\caption{Total travel numbers (2012)}
%latex.default(immigration, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{immigration}\tabularnewline
\hline
Australia&$1799655$\tabularnewline
United Kingdom&$ 401737$\tabularnewline
China&$ 322076$\tabularnewline
United States&$ 316058$\tabularnewline
Fiji&$ 151443$\tabularnewline
India&$ 107618$\tabularnewline
Japan&$ 106716$\tabularnewline
Germany&$  96308$\tabularnewline
Korea, Republic of&$  87419$\tabularnewline
France&$  85948$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:immigration12}
\end{table}

\begin{table}
\caption{Lowest national measles vaccine cover (\%, 2012)}
%latex.default(vaccinecover, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{cover}\tabularnewline
\hline
Equatorial Guinea&$34$\tabularnewline
Somalia&$49$\tabularnewline
Lesotho&$60$\tabularnewline
Central African Republic&$65$\tabularnewline
Papua New Guinea&$67$\tabularnewline
Chad&$69$\tabularnewline
Haiti&$69$\tabularnewline
South Sudan&$70$\tabularnewline
Gabon&$71$\tabularnewline
Yemen&$71$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:cover12}
\end{table}

\begin{table}
\caption{Measles incidence per million (2012)}
%latex.default(incidence, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{incidence}\tabularnewline
\hline
Equatorial Guinea&$1617$\tabularnewline
Nauru&$1100$\tabularnewline
Democratic Republic of the Congo&$1096$\tabularnewline
Somalia&$ 979$\tabularnewline
Djibouti&$ 824$\tabularnewline
Sudan&$ 786$\tabularnewline
Burkina Faso&$ 447$\tabularnewline
Romania&$ 342$\tabularnewline
Ukraine&$ 280$\tabularnewline
Sudan&$ 229$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:incidence12}
\end{table}

\begin{table}
\caption{Risk of measles importation to New Zealand in 2012}
%latex.default(risk, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{risk}\tabularnewline
\hline
Australia&$15537152$\tabularnewline
United Kingdom&$13386328$\tabularnewline
Thailand&$ 4774688$\tabularnewline
Malaysia&$ 4495338$\tabularnewline
Indonesia&$ 2216871$\tabularnewline
India&$ 1620172$\tabularnewline
China&$ 1432659$\tabularnewline
Ukraine&$  765096$\tabularnewline
Ireland&$  734128$\tabularnewline
Romania&$  655107$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:risk12}
\end{table}

\begin{table}
\caption{Immigration numbers to New Zealand (2012)}
%latex.default(immigration, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{immigration}\tabularnewline
\hline
Australia&$809775$\tabularnewline
United Kingdom&$306177$\tabularnewline
China&$256036$\tabularnewline
United States&$194438$\tabularnewline
Japan&$ 86676$\tabularnewline
Germany&$ 83608$\tabularnewline
Korea, Republic of&$ 73459$\tabularnewline
France&$ 71448$\tabularnewline
India&$ 69038$\tabularnewline
Canada&$ 54981$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:immigration12}
\end{table}

\begin{table}
\caption{Lowest national measles vaccine cover (\%, 2012)}
%latex.default(vaccinecover, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{cover}\tabularnewline
\hline
Equatorial Guinea&$34$\tabularnewline
Somalia&$49$\tabularnewline
Lesotho&$60$\tabularnewline
Central African Republic&$65$\tabularnewline
Papua New Guinea&$67$\tabularnewline
Chad&$69$\tabularnewline
Haiti&$69$\tabularnewline
South Sudan&$70$\tabularnewline
Gabon&$71$\tabularnewline
Yemen&$71$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:cover12}
\end{table}

\begin{table}
\caption{Measles incidence per million (2012)}
%latex.default(incidence, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{incidence}\tabularnewline
\hline
Equatorial Guinea&$1617$\tabularnewline
Nauru&$1100$\tabularnewline
Democratic Republic of the Congo&$1096$\tabularnewline
Somalia&$ 979$\tabularnewline
Djibouti&$ 824$\tabularnewline
Sudan&$ 786$\tabularnewline
Burkina Faso&$ 447$\tabularnewline
Romania&$ 342$\tabularnewline
Ukraine&$ 280$\tabularnewline
Sudan&$ 229$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:incidence12}
\end{table}

\begin{table}
\caption{Risk of measles importation to New Zealand in 2012}
%latex.default(risk, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{risk}\tabularnewline
\hline
United Kingdom&$10202161$\tabularnewline
Australia&$ 6991116$\tabularnewline
Malaysia&$ 3226580$\tabularnewline
Thailand&$ 1576414$\tabularnewline
China&$ 1138900$\tabularnewline
India&$ 1039356$\tabularnewline
Indonesia&$  984022$\tabularnewline
Ukraine&$  664315$\tabularnewline
Ireland&$  585413$\tabularnewline
Romania&$  490731$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:risk12}
\end{table}
%%%%%%%%%%%%%%%%%%%

\begin{table}
\caption{New Zealander travel numbers (2012)}
%latex.default(nz, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{immigration}\tabularnewline
\hline
Australia&$989880$\tabularnewline
United States&$121620$\tabularnewline
Fiji&$104720$\tabularnewline
United Kingdom&$ 95560$\tabularnewline
Cook Islands&$ 71960$\tabularnewline
China&$ 66040$\tabularnewline
Samoa&$ 46020$\tabularnewline
Thailand&$ 41100$\tabularnewline
India&$ 38580$\tabularnewline
Canada&$ 20400$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:nztravel12}
\end{table}

\begin{table}
\caption{Risk of measles importation to New Zealand due to New Zealand travellers in 2012}
%latex.default(nzrisk, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{risk}\tabularnewline
\hline
Australia&$8546036$\tabularnewline
Thailand&$3198274$\tabularnewline
United Kingdom&$3184166$\tabularnewline
Malaysia&$1268758$\tabularnewline
Indonesia&$1232849$\tabularnewline
India&$ 580816$\tabularnewline
China&$ 293759$\tabularnewline
Samoa&$ 243492$\tabularnewline
Philippines&$ 241740$\tabularnewline
Nepal&$ 188450$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:nzrisk12}
\end{table}

%%%%%%%%%%%%%%%%%%
\subsection{Risk analyses methods}
We received the raw EpiSurv measles case data from The Institute of Environmental Science and Research Ltd (ESR) on 27 June 2014. Initial analyses of those ESR data (not shown) suggested that denominator data were required to perform multivariate analyses to avoid confounding results due to a lack of independence among risk factors. Specifically we required $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data for New Zealand to test whether interactions among case covariates provide additional information on risk over the univariate analyses performed in the \emph{Progress Towards Measles Elimination in New Zealand - Final} report. These $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data were provided to us on 3 July 2014 by the University of Otago. We used these denominator data to determine if there were interactions among specific age categories, Prioritized Ethnicities, and NZDep that might exist among cases allowing better understanding of risk of measles infection.

@@

\begin{figure}[h!]
\begin{center}
\includegraphics{draftfinalreport-015}
\end{center}
\caption{Age of measles cases in years in New Zealand for two periods, 1997--2014 and 2007--2014}
\label{fig:ageinyears}
\end{figure}

@@

\begin{table}\tiny
\caption{Absolute number of measles cases in specific age, ethnicity and socio-economic deprivation categories from 2007-2014}
%latex.default(ttyr, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lllr}
\hline\hline
\multicolumn{1}{c}{NZDep}&\multicolumn{1}{c}{Age}&\multicolumn{1}{c}{Ethnicity}&\multicolumn{1}{c}{Cases}\tabularnewline
\hline
1-5&0-2&Asian&$ 11$\tabularnewline
6-10&0-2&Asian&$  8$\tabularnewline
1-5&3-5&Asian&$  1$\tabularnewline
1-5&6-17&Asian&$ 11$\tabularnewline
6-10&6-17&Asian&$  5$\tabularnewline
1-5&18-24&Asian&$  3$\tabularnewline
6-10&18-24&Asian&$  5$\tabularnewline
1-5&25+&Asian&$ 10$\tabularnewline
6-10&25+&Asian&$ 13$\tabularnewline
1-5&0-2&European&$ 83$\tabularnewline
6-10&0-2&European&$ 64$\tabularnewline
1-5&3-5&European&$ 42$\tabularnewline
6-10&3-5&European&$ 17$\tabularnewline
1-5&6-17&European&$219$\tabularnewline
6-10&6-17&European&$ 80$\tabularnewline
1-5&18-24&European&$ 34$\tabularnewline
6-10&18-24&European&$ 36$\tabularnewline
1-5&25+&European&$ 78$\tabularnewline
6-10&25+&European&$ 51$\tabularnewline
1-5&0-2&Maori&$ 18$\tabularnewline
6-10&0-2&Maori&$ 48$\tabularnewline
1-5&3-5&Maori&$  7$\tabularnewline
6-10&3-5&Maori&$ 11$\tabularnewline
1-5&6-17&Maori&$ 19$\tabularnewline
6-10&6-17&Maori&$ 92$\tabularnewline
1-5&18-24&Maori&$  5$\tabularnewline
6-10&18-24&Maori&$  8$\tabularnewline
1-5&25+&Maori&$  2$\tabularnewline
6-10&25+&Maori&$  6$\tabularnewline
6-10&0-2&MLA&$  3$\tabularnewline
6-10&3-5&MLA&$  1$\tabularnewline
6-10&6-17&MLA&$  1$\tabularnewline
6-10&18-24&MLA&$  2$\tabularnewline
6-10&25+&MLA&$  6$\tabularnewline
1-5&0-2&Pacific&$  5$\tabularnewline
6-10&0-2&Pacific&$ 58$\tabularnewline
6-10&3-5&Pacific&$  3$\tabularnewline
1-5&6-17&Pacific&$  5$\tabularnewline
6-10&6-17&Pacific&$ 22$\tabularnewline
1-5&18-24&Pacific&$  1$\tabularnewline
6-10&18-24&Pacific&$  8$\tabularnewline
1-5&25+&Pacific&$  2$\tabularnewline
6-10&25+&Pacific&$ 11$\tabularnewline
1-5&0-2&None&$  3$\tabularnewline
1-5&3-5&None&$  1$\tabularnewline
1-5&6-17&None&$  3$\tabularnewline
6-10&6-17&None&$  4$\tabularnewline
1-5&18-24&None&$  2$\tabularnewline
6-10&18-24&None&$  1$\tabularnewline
1-5&25+&None&$  5$\tabularnewline
6-10&25+&None&$  3$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:case}
\end{table}

@@

\begin{figure}[h!]
\begin{center}
\includegraphics{draftfinalreport-023}
\end{center}
\caption{Per capita rates of measles infections broken down by ethnicity, see Table~\ref{table:percap} for details}
\label{fig:percap}
\end{figure}

\begin{table}\tiny
\caption{Numbers of measles cases, population sizes and per capita rates of measles in specific age, ethnicity and socio-economic deprivation categories from 2007-2014}
%latex.default(PerCap, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lllrrr}
\hline\hline
\multicolumn{1}{c}{NZDep}&\multicolumn{1}{c}{Age}&\multicolumn{1}{c}{Ethnicity}&\multicolumn{1}{c}{Population}&\multicolumn{1}{c}{Cases 1}&\multicolumn{1}{c}{Per capita}\tabularnewline
\hline
1-5&0-2&Asian&$   6094$&$ 11$&$0.0018$\tabularnewline
6-10&0-2&Asian&$   6806$&$  8$&$0.0012$\tabularnewline
1-5&3-5&Asian&$   6094$&$  1$&$0.0002$\tabularnewline
6-10&3-5&Asian&$   6806$&$  0$&$0.0000$\tabularnewline
1-5&6-17&Asian&$  33918$&$ 11$&$0.0003$\tabularnewline
6-10&6-17&Asian&$  28905$&$  5$&$0.0002$\tabularnewline
1-5&18-24&Asian&$  22917$&$  3$&$0.0001$\tabularnewline
6-10&18-24&Asian&$  34107$&$  5$&$0.0001$\tabularnewline
1-5&25+&Asian&$  96357$&$ 10$&$0.0001$\tabularnewline
6-10&25+&Asian&$  98715$&$ 13$&$0.0001$\tabularnewline
1-5&0-2&European&$  57872$&$ 83$&$0.0014$\tabularnewline
6-10&0-2&European&$  45445$&$ 64$&$0.0014$\tabularnewline
1-5&3-5&European&$  57872$&$ 42$&$0.0007$\tabularnewline
6-10&3-5&European&$  45445$&$ 17$&$0.0004$\tabularnewline
1-5&6-17&European&$ 264330$&$219$&$0.0008$\tabularnewline
6-10&6-17&European&$ 182937$&$ 80$&$0.0004$\tabularnewline
1-5&18-24&European&$ 107649$&$ 34$&$0.0003$\tabularnewline
6-10&18-24&European&$ 117840$&$ 36$&$0.0003$\tabularnewline
1-5&25+&European&$1001916$&$ 78$&$0.0001$\tabularnewline
6-10&25+&European&$ 724317$&$ 51$&$0.0001$\tabularnewline
1-5&0-2&Maori&$  10003$&$ 18$&$0.0018$\tabularnewline
6-10&0-2&Maori&$  30104$&$ 48$&$0.0016$\tabularnewline
1-5&3-5&Maori&$  10003$&$  7$&$0.0007$\tabularnewline
6-10&3-5&Maori&$  30104$&$ 11$&$0.0004$\tabularnewline
1-5&6-17&Maori&$  40461$&$ 19$&$0.0005$\tabularnewline
6-10&6-17&Maori&$ 116640$&$ 92$&$0.0008$\tabularnewline
1-5&18-24&Maori&$  15360$&$  5$&$0.0003$\tabularnewline
6-10&18-24&Maori&$  48495$&$  8$&$0.0002$\tabularnewline
1-5&25+&Maori&$  71217$&$  2$&$0.0000$\tabularnewline
6-10&25+&Maori&$ 192729$&$  6$&$0.0000$\tabularnewline
1-5&0-2&MLA&$    728$&$  0$&$0.0000$\tabularnewline
6-10&0-2&MLA&$   1290$&$  3$&$0.0023$\tabularnewline
1-5&3-5&MLA&$    728$&$  0$&$0.0000$\tabularnewline
6-10&3-5&MLA&$   1290$&$  1$&$0.0008$\tabularnewline
1-5&6-17&MLA&$   2991$&$  0$&$0.0000$\tabularnewline
6-10&6-17&MLA&$   4539$&$  1$&$0.0002$\tabularnewline
1-5&18-24&MLA&$   1710$&$  0$&$0.0000$\tabularnewline
6-10&18-24&MLA&$   3078$&$  2$&$0.0006$\tabularnewline
1-5&25+&MLA&$   8028$&$  0$&$0.0000$\tabularnewline
6-10&25+&MLA&$  10335$&$  6$&$0.0006$\tabularnewline
1-5&0-2&Pacific&$   2093$&$  5$&$0.0024$\tabularnewline
6-10&0-2&Pacific&$  13124$&$ 58$&$0.0044$\tabularnewline
1-5&3-5&Pacific&$   2093$&$  0$&$0.0000$\tabularnewline
6-10&3-5&Pacific&$  13124$&$  3$&$0.0002$\tabularnewline
1-5&6-17&Pacific&$   8541$&$  5$&$0.0006$\tabularnewline
6-10&6-17&Pacific&$  51183$&$ 22$&$0.0004$\tabularnewline
1-5&18-24&Pacific&$   3972$&$  1$&$0.0003$\tabularnewline
6-10&18-24&Pacific&$  22098$&$  8$&$0.0004$\tabularnewline
1-5&25+&Pacific&$  18492$&$  2$&$0.0001$\tabularnewline
6-10&25+&Pacific&$  91533$&$ 11$&$0.0001$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:percap}
\end{table}

@@

>>>>

\section{Additional risk analyses}
\label{sub:risk_analyses}

In this section we provide work that we believe will help inform the Ministry of Health regarding the understanding of risk from measles. These analyses are intended to build on the analyses already included in the \emph {Progress Towards Measles Elimination in New Zealand - Final} report reviewed above.

We include multivariate modelling to account for confounding within the univariate analyses for measles cases in New Zealand, and descriptive analyses of risk of infection due to previous vaccination history, current rates of immunity within the population. In light of the apparent increasing trend in measles incidence in the last few years (Figure~\ref{fig:incidence1997}), we reviewed the information on measles importation and the origins of the introductions of measles into New Zealand. To help understand the risk of measles importation, with a particular goal of enabling the Ministry to better inform travellers and understand high risk periods, we sought to quantitatively evaluate the risk of measles importation from travel.

@@

%%%%%%%%%%%%%%%%%%
\subsection{Risk of measles infection in New Zealand analyses methods}
We received the raw EpiSurv measles case data from The Institute of Environmental Science and Research Ltd (ESR) on 27 June 2014. Initial analyses of those ESR data (not shown) suggested that denominator data were required to perform multivariate analyses to avoid confounding results due to a lack of independence among risk factors. Specifically we required $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data for New Zealand to test whether interactions among case covariates provide additional information on risk over the univariate analyses performed in the \emph{Progress Towards Measles Elimination in New Zealand - Final} report. These $\textsf{Age} \times \textsf{Prioritised Ethnicity} \times \textsf{NZDep}$ data were provided to us on 3 July 2014 by the University of Otago. We used these denominator data to determine if there were interactions among specific age categories, Prioritized Ethnicities, and NZDep that might exist among cases allowing better understanding of risk of measles infection.

@@

\begin{figure}
\begin{center}
\includegraphics{draftfinalreport-002}
\end{center}
\caption{Age of measles cases in years in New Zealand for two periods, 1997--2014 and 2007--2014}
\label{fig:ageinyears}
\end{figure}

@@

\begin{table}[hbtp]
\footnotesize
%latex.default(ttyr, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lllr}
\hline\hline
\multicolumn{1}{c}{NZDep}&\multicolumn{1}{c}{Age}&\multicolumn{1}{c}{Ethnicity}&\multicolumn{1}{c}{Cases}\tabularnewline
\hline
1-5&0-2&Asian&$ 11$\tabularnewline
6-10&0-2&Asian&$  8$\tabularnewline
1-5&3-5&Asian&$  1$\tabularnewline
1-5&6-17&Asian&$ 11$\tabularnewline
6-10&6-17&Asian&$  5$\tabularnewline
1-5&18-24&Asian&$  3$\tabularnewline
6-10&18-24&Asian&$  5$\tabularnewline
1-5&25+&Asian&$ 10$\tabularnewline
6-10&25+&Asian&$ 13$\tabularnewline
1-5&0-2&European&$ 83$\tabularnewline
6-10&0-2&European&$ 64$\tabularnewline
1-5&3-5&European&$ 42$\tabularnewline
6-10&3-5&European&$ 17$\tabularnewline
1-5&6-17&European&$219$\tabularnewline
6-10&6-17&European&$ 80$\tabularnewline
1-5&18-24&European&$ 34$\tabularnewline
6-10&18-24&European&$ 36$\tabularnewline
1-5&25+&European&$ 78$\tabularnewline
6-10&25+&European&$ 51$\tabularnewline
1-5&0-2&Maori&$ 18$\tabularnewline
6-10&0-2&Maori&$ 48$\tabularnewline
1-5&3-5&Maori&$  7$\tabularnewline
6-10&3-5&Maori&$ 11$\tabularnewline
1-5&6-17&Maori&$ 19$\tabularnewline
6-10&6-17&Maori&$ 92$\tabularnewline
1-5&18-24&Maori&$  5$\tabularnewline
6-10&18-24&Maori&$  8$\tabularnewline
1-5&25+&Maori&$  2$\tabularnewline
6-10&25+&Maori&$  6$\tabularnewline
6-10&0-2&MLA&$  3$\tabularnewline
6-10&3-5&MLA&$  1$\tabularnewline
6-10&6-17&MLA&$  1$\tabularnewline
6-10&18-24&MLA&$  2$\tabularnewline
6-10&25+&MLA&$  6$\tabularnewline
1-5&0-2&Pacific&$  5$\tabularnewline
6-10&0-2&Pacific&$ 58$\tabularnewline
6-10&3-5&Pacific&$  3$\tabularnewline
1-5&6-17&Pacific&$  5$\tabularnewline
6-10&6-17&Pacific&$ 22$\tabularnewline
1-5&18-24&Pacific&$  1$\tabularnewline
6-10&18-24&Pacific&$  8$\tabularnewline
1-5&25+&Pacific&$  2$\tabularnewline
6-10&25+&Pacific&$ 11$\tabularnewline
1-5&0-2&None&$  3$\tabularnewline
1-5&3-5&None&$  1$\tabularnewline
1-5&6-17&None&$  3$\tabularnewline
6-10&6-17&None&$  4$\tabularnewline
1-5&18-24&None&$  2$\tabularnewline
6-10&18-24&None&$  1$\tabularnewline
1-5&25+&None&$  5$\tabularnewline
6-10&25+&None&$  3$\tabularnewline
\hline
\end{tabular}\end{center}\caption{Absolute number of measles cases in specific age, ethnicity and socio-economic deprivation categories from 2007-2014}
\label{table:case}
\end{table}

@@

\begin{figure}[h!]
\begin{center}
\includegraphics{draftfinalreport-010}
\end{center}
\caption{Per capita rates of measles infections broken down by ethnicity, see Table~\ref{table:percap} for details}
\label{fig:percap}
\end{figure}

\begin{table}[hbtp]
\footnotesize
\caption{Numbers of measles cases, population sizes and per capita rates of measles in specific age, ethnicity and socio-economic deprivation categories from 2007-2014}
%latex.default(PerCap, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lllrrr}
\hline\hline
\multicolumn{1}{c}{NZDep}&\multicolumn{1}{c}{Age}&\multicolumn{1}{c}{Ethnicity}&\multicolumn{1}{c}{Population}&\multicolumn{1}{c}{Cases 1}&\multicolumn{1}{c}{Per capita}\tabularnewline
\hline
1-5&0-2&Asian&$   6094$&$ 11$&$0.0018$\tabularnewline
6-10&0-2&Asian&$   6806$&$  8$&$0.0012$\tabularnewline
1-5&3-5&Asian&$   6094$&$  1$&$0.0002$\tabularnewline
6-10&3-5&Asian&$   6806$&$  0$&$0.0000$\tabularnewline
1-5&6-17&Asian&$  33918$&$ 11$&$0.0003$\tabularnewline
6-10&6-17&Asian&$  28905$&$  5$&$0.0002$\tabularnewline
1-5&18-24&Asian&$  22917$&$  3$&$0.0001$\tabularnewline
6-10&18-24&Asian&$  34107$&$  5$&$0.0001$\tabularnewline
1-5&25+&Asian&$  96357$&$ 10$&$0.0001$\tabularnewline
6-10&25+&Asian&$  98715$&$ 13$&$0.0001$\tabularnewline
1-5&0-2&European&$  57872$&$ 83$&$0.0014$\tabularnewline
6-10&0-2&European&$  45445$&$ 64$&$0.0014$\tabularnewline
1-5&3-5&European&$  57872$&$ 42$&$0.0007$\tabularnewline
6-10&3-5&European&$  45445$&$ 17$&$0.0004$\tabularnewline
1-5&6-17&European&$ 264330$&$219$&$0.0008$\tabularnewline
6-10&6-17&European&$ 182937$&$ 80$&$0.0004$\tabularnewline
1-5&18-24&European&$ 107649$&$ 34$&$0.0003$\tabularnewline
6-10&18-24&European&$ 117840$&$ 36$&$0.0003$\tabularnewline
1-5&25+&European&$1001916$&$ 78$&$0.0001$\tabularnewline
6-10&25+&European&$ 724317$&$ 51$&$0.0001$\tabularnewline
1-5&0-2&Maori&$  10003$&$ 18$&$0.0018$\tabularnewline
6-10&0-2&Maori&$  30104$&$ 48$&$0.0016$\tabularnewline
1-5&3-5&Maori&$  10003$&$  7$&$0.0007$\tabularnewline
6-10&3-5&Maori&$  30104$&$ 11$&$0.0004$\tabularnewline
1-5&6-17&Maori&$  40461$&$ 19$&$0.0005$\tabularnewline
6-10&6-17&Maori&$ 116640$&$ 92$&$0.0008$\tabularnewline
1-5&18-24&Maori&$  15360$&$  5$&$0.0003$\tabularnewline
6-10&18-24&Maori&$  48495$&$  8$&$0.0002$\tabularnewline
1-5&25+&Maori&$  71217$&$  2$&$0.0000$\tabularnewline
6-10&25+&Maori&$ 192729$&$  6$&$0.0000$\tabularnewline
1-5&0-2&MLA&$    728$&$  0$&$0.0000$\tabularnewline
6-10&0-2&MLA&$   1290$&$  3$&$0.0023$\tabularnewline
1-5&3-5&MLA&$    728$&$  0$&$0.0000$\tabularnewline
6-10&3-5&MLA&$   1290$&$  1$&$0.0008$\tabularnewline
1-5&6-17&MLA&$   2991$&$  0$&$0.0000$\tabularnewline
6-10&6-17&MLA&$   4539$&$  1$&$0.0002$\tabularnewline
1-5&18-24&MLA&$   1710$&$  0$&$0.0000$\tabularnewline
6-10&18-24&MLA&$   3078$&$  2$&$0.0006$\tabularnewline
1-5&25+&MLA&$   8028$&$  0$&$0.0000$\tabularnewline
6-10&25+&MLA&$  10335$&$  6$&$0.0006$\tabularnewline
1-5&0-2&Pacific&$   2093$&$  5$&$0.0024$\tabularnewline
6-10&0-2&Pacific&$  13124$&$ 58$&$0.0044$\tabularnewline
1-5&3-5&Pacific&$   2093$&$  0$&$0.0000$\tabularnewline
6-10&3-5&Pacific&$  13124$&$  3$&$0.0002$\tabularnewline
1-5&6-17&Pacific&$   8541$&$  5$&$0.0006$\tabularnewline
6-10&6-17&Pacific&$  51183$&$ 22$&$0.0004$\tabularnewline
1-5&18-24&Pacific&$   3972$&$  1$&$0.0003$\tabularnewline
6-10&18-24&Pacific&$  22098$&$  8$&$0.0004$\tabularnewline
1-5&25+&Pacific&$  18492$&$  2$&$0.0001$\tabularnewline
6-10&25+&Pacific&$  91533$&$ 11$&$0.0001$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:percap}
\end{table}

@@

\subsection{Measles importation risk}

For our measles importation risk analyses, we use arrivals data from New Zealand immigration and New Zealander travel data by country and year (\href{http://www.immigration.govt.nz/}{www.immigration.govt.nz}) to measure human movement to and from New Zealand. We collated country population size, measles incidence and measles vaccination cover from the WHO (\href{http://www.who.int/research/en/}{www.who.int/research/en/}). Note the immigration figures use all immigration of foreign nationals, coming for whatever purpose and includes non-New Zealanders resident in New Zealand, but not yet holders of New Zealand passports. We used the WHO data to determine per capita measles cases for each year and used these data and the number of immigrants to New Zealand to begin to understand where measles was likely to be imported from. We used simple per capita rates for measles and the number of travellers to each country to score and map the risk of mealses importation. We use the data from 2012 because this year had the most complete WHO measles data and yet was most recent, thus accounts for improved measles vaccination coverage following the United Nations Millenium Development Goals' improvements in measles vaccination coverage.

\begin{table}
\caption{Lowest national measles vaccine cover (\%, 2012)}
%latex.default(vaccinecover, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{cover}\tabularnewline
\hline
Equatorial Guinea&$34$\tabularnewline
Somalia&$49$\tabularnewline
Lesotho&$60$\tabularnewline
Central African Republic&$65$\tabularnewline
Papua New Guinea&$67$\tabularnewline
Chad&$69$\tabularnewline
Haiti&$69$\tabularnewline
South Sudan&$70$\tabularnewline
Gabon&$71$\tabularnewline
Yemen&$71$\tabularnewline
Benin&$72$\tabularnewline
Lao People's Democratic Republic&$72$\tabularnewline
Togo&$72$\tabularnewline
Suriname&$73$\tabularnewline
Timor-Leste&$73$\tabularnewline
Paraguay&$74$\tabularnewline
Mauritania&$75$\tabularnewline
Namibia&$76$\tabularnewline
Eritrea&$77$\tabularnewline
Marshall Islands&$78$\tabularnewline
Nigeria&$78$\tabularnewline
Syrian Arab Republic&$78$\tabularnewline
Ukraine&$79$\tabularnewline
Congo&$80$\tabularnewline
Ethiopia&$80$\tabularnewline
Liberia&$80$\tabularnewline
Afghanistan&$81$\tabularnewline
Cameroon&$82$\tabularnewline
Mozambique&$82$\tabularnewline
Senegal&$82$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:cover12}
\end{table}

\begin{table}
\caption{Measles incidence per million (2012)}
%latex.default(incidence, file = "", table.env = FALSE, rowname = NULL)%
\begin{center}
\begin{tabular}{lr}
\hline\hline
\multicolumn{1}{c}{country}&\multicolumn{1}{c}{incidence}\tabularnewline
\hline
Equatorial Guinea&$1617$\tabularnewline
Nauru&$1100$\tabularnewline
Democratic Republic of the Congo&$1096$\tabularnewline
Somalia&$ 979$\tabularnewline
Djibouti&$ 824$\tabularnewline
Sudan&$ 786$\tabularnewline
Burkina Faso&$ 447$\tabularnewline
Romania&$ 342$\tabularnewline
Ukraine&$ 280$\tabularnewline
Sudan&$ 229$\tabularnewline
Angola&$ 214$\tabularnewline
Monaco&$ 132$\tabularnewline
Nepal&$ 122$\tabularnewline
Sierra Leone&$ 113$\tabularnewline
Afghanistan&$  93$\tabularnewline
Yemen&$  91$\tabularnewline
Lesotho&$  87$\tabularnewline
Qatar&$  78$\tabularnewline
Thailand&$  78$\tabularnewline
Malaysia&$  64$\tabularnewline
Zambia&$  64$\tabularnewline
Ghana&$  64$\tabularnewline
Indonesia&$  63$\tabularnewline
Congo&$  60$\tabularnewline
Uganda&$  56$\tabularnewline
Libyan Arab Jamahiriya&$  52$\tabularnewline
Ethiopia&$  47$\tabularnewline
Pakistan&$  45$\tabularnewline
Myanmar&$  41$\tabularnewline
Nigeria&$  38$\tabularnewline
\hline
\end{tabular}\end{center}\label{table:incidence12}
\end{table}
